---
title: "Covid91 vaccine study Final2024"
author: "Put your name here!"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

$H_0$: The Covid921 vaccine is not effective in MALES  

$H_A$: The Covid921 vaccine is effective in MALES 

### Methods 

Treatment is a categorical variable  
MALES is also a categorical variable  
So we use the cat~cat helper file  

### Descriptive Results 

### Graphical Results

```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected,data=males)
barchartGC(~infected + treatment,data=males, type="percent")
```

The barchart shows that MALES who got the vaccines were less likely to get Covid921, than those who got the placebo. However, about 50% of MALES that got vaccine were OK, and about 50% of MALES with the placebo were OK.

### Numerical Results

```{r}
table1 <- xtabs(~treatment + infected,data=males)
rowPerc(table1)
colPerc(table1)
```


The numerical results show that about 74% of MALES who took the placebo got Covid921, while about 26% of MALES who got the vaccines got Covid921.

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi-squared test shows that the p-value = 0.00000000000000022, so we reject the null hypothesis since it is less than 0.05.

The Fisher's Exact Test shows that the p-value is 0.00000000000000022, so we reject the null hypothesis since it is less than 0.05. The test also shows that MALES have an odds ratio of being 2.8 times more likely to get Covid921 with a placebo compared to those who got the vaccine. 

## Females

$H_0$: The Covid921 vaccine is not effective in FEMALES 

$H_A$: The Covid921 vaccine is effective in FEMALES

### Methods 

Treatment is a categorical variable   
FEMALES is a categorical variable   
So we use is cat~cat helper file 

### Descriptive Results 

### Graphical Results

```{r}
females=filter(FinalData,FinalData$sex=="F")
barchartGC(~treatment + infected,data=females)
barchartGC(~infected + treatment,data=females, type="percent")
```

The barchart shows that FEMALES who got the vaccines were less likely to get Covid921, than those who got the placebo. However, about 50% of FEMALES that got the vaccine were OK, and about 50% of FEMALES with the placebo were OK.

### Numerical Results

```{r}
table1 <- xtabs(~treatment + infected,data=females)
rowPerc(table1)
colPerc(table1)
```

The numerical results data indicates that the Covid921 vaccine significantly decreases the chances of contracting the virus compared to the placebo. Specifically, it reduces the odds of getting Covid921 by about 46.9% and lowers the risk of infection by approximately 45.5% compared to the placebo.

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi-squared test shows that the p-value = 0.00000001914, so we reject the null hypothesis because it is less than 0.05.

The Fisher's Exact Test shows that the p-value = 0.00000001246, so we reject the null hypothesis because it is less than 0.05. The test also shows that FEMALES have a odds ratio of being 1.86 times more likely to get Covid921 with a placebo compared to those who got the vaccine.  

## LGBTQ

$H_0$: The Covid921 vaccine is not effective in LGBTQ Community.

$H_A$: The Covid921 vaccine is effective in LGBTQ Community.

### Methods 

Treatment is a categorical variable   
LGBTQ is a categorical variable   
So we use is cat~cat helper file    

### Descriptive Results 

### Graphical Results

```{r}
LGBTQ = filter(FinalData, LGBTQ=="gay")

barchartGC(~infected + treatment,data=LGBTQ)
barchartGC(~infected + treatment,data=LGBTQ, type="percent")
```

The barchart shows that LGBTQ who got the vaccines were less likely to get Covid921, than those who got the placebo. However, about 50% of LGBTQ that got the vaccine were OK, and about 45% of LGBTQ with the placebo were OK.

### Numerical Results

```{r}
table1 <- xtabs(~treatment + infected, data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```

The numerical results show that 80% of LGBTQ participants received a placebo and 20% received the Covid921 vaccine. The Covid921 vaccine reduced the odds of contracting the virus to approximately 24% compared to the placebo. The risk of infection with the vaccine was about 44.8% of the risk with the placebo, resulting in a relative risk reduction of around 55.2%, indicating a significant protective effect.

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi-squared test shows that the p-value = 0.000002266 which is smaller than 0.05. This shows that the vaccine does significantly reduce the risk of contracting Covid921 among the LGBTQ community, so we reject the null hypothesis.

The Fisher's Exact Test p-value = 0.0000007269, and the odds of getting Covid921 with the vaccine is about 0.24 times the odds of getting it with the placebo. Suggesting that the vaccine considerably lowers the risk compared to the placebo.

## Druggies

$H_0$: The Covid921 vaccine is not effective in Drug users 

$H_A$: The Covid921 vaccine is effective in Drug users

### Methods 

Treatment is a categorical variable   
DRUGGIES is a categorical variable   
So we use is cat~cat helper file    

### Descriptive Results 

### Graphical Results

```{r}
DrugUser=filter(FinalData,FinalData$DrugUser=="yes")

barchartGC(~treatment + infected,data=DrugUser)
barchartGC(~infected + treatment,data=DrugUser,type="percent")
```

The barchart shows that the Drug Users that got the placebo were less likely to get Covid921 then those who got the vaccine. However, both were about 50% OK.

# Overall Results and Conclusions