---
title: "Covid91 vaccine study Final2024"
author: "La Kayla Thompson"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

$H_0$: The Covid921 vaccine is not effective in MALES  

$H_A$: The Covid921 vaccine is effective in MALES 

### Methods 

Treatment is a categorical variable  
MALES is also a categorical variable  
So we use the cat~cat helper file  

### Descriptive Results 

### Graphical Results

```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected,data=males)
barchartGC(~infected + treatment,data=males, type="percent")
```

The bar charts indicate that MALES who received the vaccine were less likely to contract Covid921 compared to those who received a placebo. However, about 50% of MALES who received either the vaccine or placebo did not contract the virus.

### Numerical Results

```{r}
table1 <- xtabs(~treatment + infected,data=males)
rowPerc(table1)
colPerc(table1)
```


The numerical results show that about 74% of MALES who received the placebo contracted Covid921, whereas only about 26% of MALES who received the vaccine contracted the virus.

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi-squared test and Fisher's Exact Test both show a p-value of 0.00000000000000022, which is less than 0.05. Therefore, we reject the null hypothesis, indicating a statistically significant effect of the vaccine. MALES are 2.8 times more likely to contract Covid921 with the placebo compared to the vaccine.

## Females

$H_0$: The Covid921 vaccine is not effective in FEMALES 

$H_A$: The Covid921 vaccine is effective in FEMALES

### Methods 

Treatment is a categorical variable   
FEMALES is a categorical variable   
So we use is cat~cat helper file 

### Descriptive Results 

### Graphical Results

```{r}
females=filter(FinalData,FinalData$sex=="F")
barchartGC(~treatment + infected,data=females)
barchartGC(~infected + treatment,data=females, type="percent")
```

The bar charts shows that FEMALES who received the vaccine were less likely to contract Covid921 compared to those who received the placebo. However, about 50% of FEMALES in both groups remained uninfected.

### Numerical Results

```{r}
table1 <- xtabs(~treatment + infected,data=females)
rowPerc(table1)
colPerc(table1)
```

The numerical results shows that the Covid921 vaccine reduces the odds of contracting the virus by approximately 46.9% and lowers the risk of infection by around 45.5% compared to the placebo.

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi-squared test yields a p-value of 0.00000001914, and the Fisher's Exact Test shows a p-value of 0.00000001246, both less than 0.05. Thus, we reject the null hypothesis. The odds ratio indicates that FEMALES are 1.86 times more likely to contract Covid921 with the placebo than with the vaccine.

## LGBTQ

$H_0$: The Covid921 vaccine is not effective in LGBTQ Community.

$H_A$: The Covid921 vaccine is effective in LGBTQ Community.

### Methods 

Treatment is a categorical variable   
LGBTQ is a categorical variable   
So we use is cat~cat helper file    

### Descriptive Results 

### Graphical Results

```{r}
LGBTQ = filter(FinalData, LGBTQ=="gay")

barchartGC(~infected + treatment,data=LGBTQ)
barchartGC(~infected + treatment,data=LGBTQ, type="percent")
```

The bar charts show that LGBTQ individuals who received the vaccine were less likely to contract Covid921 compared to those who received the placebo. However, about 45% of LGBTQ that got the placebo were uninfected, and 50% that got the vaccine were uninfected.

### Numerical Results

```{r}
table1 <- xtabs(~treatment + infected, data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```

The numerical results show that about 80% of LGBTQ participants who received the placebo contracted Covid921, while 20% who received the vaccine contracted the virus. The vaccine reduces the relative risk of infection by approximately 55.2%.

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi-squared test gives a p-value of 0.000002266, and the Fisher's Exact Test provides a p-value of 0.0000007269, both below 0.05, so we reject the null hypothesis. This indicates a statistically significant effect of the vaccine. The odds of contracting Covid921 are significantly lower with the vaccine compared to the placebo.

## Druggies

$H_0$: The Covid921 vaccine is not effective in Drug users 

$H_A$: The Covid921 vaccine is effective in Drug users

### Methods 

Treatment is a categorical variable   
DRUGGIES is a categorical variable   
So we use is cat~cat helper file    

### Descriptive Results 

### Graphical Results

```{r}
DrugUser=filter(FinalData,FinalData$DrugUser=="yes")

barchartGC(~treatment + infected,data=DrugUser)
barchartGC(~infected + treatment,data=DrugUser,type="percent")
```

The bar charts shows an unexpected result: DRUG USERS who received the placebo were less likely to contract Covid921 compared to those who received the vaccine.  About 50% of DRUG USERS in both groups remained uninfected. 

### Numerical Results

```{r}
table1 <- xtabs(~treatment + infected, data=DrugUser)
rowPerc(table1)
colPerc(table1)
```

The numerical results a 3.28% difference in infection rates: 3.61% of those who received the vaccine contracted Covid921 compared to 0.33% of those who received the placebo. This suggests the vaccine may not be effective in DRUG USERS.Typically, a vaccine is expected to reduce the risk of Covid921, not increase it.

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi-squared test yields a p-value of 0.0000000001414, and the Fisher's Exact Test provides a p-value of 0.000000000003261. Both p-values are less than 0.05, indicating a significant difference between the vaccine and placebo groups. Statistically we would reject the null hypothesis. However, the data suggests an increased risk of infection for DRUG USERS who received the vaccine, which is an unexpected finding. Therefore, we fail to reject the null hypothesis.

# Individual Conclusions

## MALES

In the study, the data showed that MALES who received the vaccine were significantly less likely to contract Covid921, with only 26% infection among the vaccinated compared to 74% among those who received a placebo. Inferential analysis using the chi-squared and Fisher's Exact Test yielded extremely low p-values (0.00000000000000022), well below 0.05, leading to the rejection of the null hypothesis. Thus, the vaccine is concluded to be effective in reducing the risk of contracting Covid921 in MALES.

## FEMALES

The study on FEMALES, those who received the vaccine had a significantly lower incidence of Covid921 compared to those who received a placebo. The vaccine reduced the odds of infection by approximately 46.9%. Inferential tests,  chi-squared and Fisher's Exact Test, yielded p-values significantly below 0.05 (0.00000001914 and 0.00000001246, respectively), leading to the rejection of the null hypothesis. This strongly supports that the Covid921 vaccine is effective in reducing infections among FEMALES.

## LGBTQ

IN the study for the LGBTQ community,the graphical and numerical results demonstrated a significant reduction in the incidence of Covid921 among vaccine recipients. Showing a relative risk reduction of approximately 55.2% compared to the placebo group. The inferential results, supported by extremely low p-values from the chi-squared (0.000002266) and Fisher's Exact Test (0.0000007269), led to the rejection of the null hypothesis. This indicates that the vaccine significantly reduces the risk of contracting Covid921 in the LGBTQ community.

## DRUG USERS

In the case of DRUG USERS, the graphical and numerical analysis revealed a higher infection rate in the vaccine group (3.61%) compared to the placebo group (0.33%), contrary to expectations. Despite these unexpected results, both the chi-squared and Fisher's Exact Test yielded extremely significant p-values (0.0000000001414 and 0.000000000003261, respectively). This suggests a statistically significant difference between the groups, but the observed increased infection risk with the vaccine requires further investigation to clarify these anomalous findings. With this information with the increased infection risk with the vaccine leads us to into failing to reject the null hypothesis.

# Overall Results and Conclusion

The Covid921 vaccine is shown to be significantly effective in reducing the cases of Covid921 among MALES, FEMALES, and LGBTQ communities, as the null hypothesis was consistently rejected with extremely low p-values, indicating a strong protective effect compared to a placebo. However, the data for DRUG USERS suggests an unexpected increase in infection risk regardless of having a low p-value as well. This highlights the need for further investigation into possible data anomalies or specific contextual factors. Overall, the vaccine is recommended for the general population, but caution and additional analysis are advised for DRUG USERS.

